
The process of DNA extraction is as follows:

1. Take a water sample (~100ml) from a boat hold
2. Centrifuge to seperate DNA and place in sample tube (~5ml)
3. Split sample into 3 'reactions' (~1ml)
4. Add Master mix (either DORY primer or DLOOP primer) - for DNA sequence of known length (e.g., 2140 base pairs)
5. Weight total DNA in replicate
6. PCR (95C, ~45C, 72C) for ~40 cycles - check the RFU for the cycles
7. Gel electrophoresis to make sure the sequence of interest is corrected amplified (does primer work?) Should only be peak at 2.140K bp length on gel (even if different species, the base pairs should only be a couple of BPs out)
8. Combine the three replicates back into the single sample
9. Add universal barcode to each sample (to then identify samples later)
10. Combine samples and put thorugh nanopore to sequence the DNA
11. Use the barcode to split back to the samples 
12. Use clustering algorithm (CD-HIT) to put into most reasonable clusters (the hope is that each of these correspond to a single OTU)
13. Use MEDAKA algorithm to get 'average' sequence for each cluster 
14. Search OTU in the NCBI database
15. If two clusters correspond to the same species then combine the number of reads. 


```{r}
pak::pak("bruigtp/flowchart")
library(flowchart)
```


```{r}
safo
```